---
title: "8160 data gen"
date: "2023-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(glmnet)
library(tidyverse)
library(ggplot2)
```

```{r data generation function 1}
data_gen1 <- function(n_beta = 40,
                     n_strong = 5,
                     n_wai = 10,
                     n_wbc = 10,
                     cor = 0.3,
                     n_samps = 10000,
                     c = 3){
  # compute the number of null predictors to generate
  n_null = n_beta - n_strong - n_wai - n_wbc
  
  # randomly generate site for strong, wai, and wbc predictors
  b_strong_ind <- sample(seq(1, n_beta), size = n_strong) 
  b_wai_ind <- sample(seq(1, n_beta)[-b_strong_ind], size = n_wai)
  b_wbc_ind <- sample(seq(1, n_beta)[-c(b_strong_ind, b_wai_ind)], size = n_wbc)
  b_null_ind <- sample(seq(1, n_beta)[-c(b_strong_ind, b_wai_ind, b_wbc_ind)],
                       size = n_null)
  
  # generate correlation matrix
  cor_mat <- diag(n_beta)
  cor_mat[b_strong_ind, b_wbc_ind] <- cor
  cor_mat[b_wbc_ind, b_strong_ind] <- cor
  
  # generate data from multivariate normal distribution from the 
  # above correlation matrix
  X = mvrnorm(n = n_samps, mu = rep(0, n_beta), 
            Sigma = cor_mat, 
            empirical = F,
            tol = 0.9)
  
  # threshold for strong vs. weak beta coefficients
  threshold = c*sqrt(log(n_beta)/n_samps)
  
  # generate betas
  strong_beta <- runif(n_strong, threshold, 3*threshold) # strong betas
  wai_beta <- runif(n_wai, 0, threshold) # wai betas
  wbc_beta <- runif(n_wbc, 0, threshold) # wbc betas
  
  # assign betas based on predictor ndex
  all_beta <- rep(0, n_beta)
  all_beta[b_strong_ind] <- strong_beta
  all_beta[b_wai_ind] <- wai_beta
  all_beta[b_wbc_ind] <- wbc_beta
  
  # generate data where forward selection and LASSO are to be ran
  Y = 1 + X %*% all_beta + rnorm(n_samps)
  df <- data.frame(cbind(X, Y))
  names(df)[n_beta + 1] <- "y"
  return(list(generated_data = df,
              b_strong_ind = b_strong_ind,
              b_wai_ind = b_wai_ind,
              b_wbc_ind = b_wbc_ind,
              b_null_ind = b_null_ind,
              beta_0=all_beta))
}

data1 <- data_gen1()

```

Possible scenarios for simulation: (1) Change the # of WBC, WAI, NULL predictors; (2) Add correlation between null predictors and strong predictors

```{r forward selection function}
fit_forward <- function(data_gen_obj){
  
  # extract generated data 
  df <- data_gen_obj$generated_data 
  
  # run forward selection
  fit <- step(object = lm(y ~ 1, data = df),
                      scope = formula(lm(y ~ ., data = df)), 
                      direction = "forward", 
                      k = 2, 
                      trace = 0) 
  # extract information for later use
  forward_summ <- summary(fit) # fitted model summary
  var_selected <- variable.names(fit)[-1] # selected variables
  
  return(list(est_beta = forward_summ$coefficient[,1][-1],
              var_selected = var_selected))
}

forward_summ <- fit_forward(data1)
forward_summ
```

```{r lasso function}
fit_lasso <- function(data_gen_obj){
  
  # extract generated data 
  df <- data_gen_obj$generated_data 
  
  # run lasso
  fit <- cv.glmnet(as.matrix(df[,-ncol(df)]),
                       df$y, 
                       nfolds = 5, 
                       type.measure = "mse") # 5-fold CV using mean square error
  param_best <- fit$glmnet.fit$beta[, fit$lambda == fit$lambda.1se] 
  
  # extract information for later use
  est_beta <- param_best[param_best != 0]
  var_selected <- names(est_beta) # selected variables
  
  return(list(var_selected = var_selected,
              est_beta = est_beta))
}

lasso_summ <- fit_lasso(data1)
lasso_summ
```

```{r calculate beta_se functions}
check_beta_se<-function(data_obj,fit_obj){
  #get beta_0
  beta_0<-data_obj$beta_0
  #get beta_hat
  var_select<-names(fit_obj$est_beta)
  beta_select<-unname(fit_obj$est_beta)

  new_vec <- as.numeric(gsub("X", "",var_select))

  beta_hat=rep(0,length(beta_0))
  beta_hat[new_vec]=unname(fit_obj$est_beta) 
  
  #
  beta_data<-data.frame(beta_0,beta_hat) %>% 
    mutate(se=(beta_0-beta_hat)^2)
  return(list(beta_data$se,sse=sum(beta_data$se)))
}

check_beta_se(data1,forward_summ)
check_beta_se(data1,lasso_summ)
```




```{r helper functions}
# check whether strong/wbc/wai indicators were correctly identified
check_pct_identified <- function(data_obj, fit_obj){
  # get indices
  b_strong_ind <- data_obj$b_strong_ind
  b_wbc_ind <- data_obj$b_wbc_ind
  b_wai_ind <- data_obj$b_wai_ind
  b_null_ind <- data_obj$b_null_ind
  
  # get variable names
  strong_ind_name <- paste("X", b_strong_ind, sep="")
  wai_ind_name <- paste("X", b_wai_ind, sep="")
  wbc_ind_name <- paste("X", b_wbc_ind, sep="")
  null_ind_name <- paste("X", b_null_ind, sep = "")
  predictor_ind_name <- c(strong_ind_name, wai_ind_name, wbc_ind_name)
  
  # check how many strong/wbc/wai/null indicators were selected
  pct_strong_identified <- length(intersect(strong_ind_name, fit_obj$var_selected))/length(strong_ind_name) * 100
  pct_wai_identified <- length(intersect(wai_ind_name, fit_obj$var_selected))/length(wai_ind_name) * 100
  pct_wbc_identified <- length(intersect(wbc_ind_name, fit_obj$var_selected))/length(wbc_ind_name) * 100
  pct_null_identified <- length(intersect(null_ind_name, fit_obj$var_selected))/length(null_ind_name) * 100 # type 1 error
  pct_identified <- length(intersect(predictor_ind_name, fit_obj$var_selected))/length(predictor_ind_name) * 100 # power
  pct_weak_identified <- length(intersect(c(wbc_ind_name, wai_ind_name), fit_obj$var_selected))/length(c(wbc_ind_name, wai_ind_name)) * 100

  return(list(pct_strong_identified = pct_strong_identified,
              pct_wai_identified = pct_wai_identified,
              pct_wbc_identified = pct_wbc_identified,
              type1_error = pct_null_identified,
              power = pct_identified,
              pct_weak_identified = pct_weak_identified))
}

check_pct_identified(data1, lasso_summ)
```


## Varying correlation (0.1, 0.3, 0.5, 0.7, 0.9)
### Forward selection
Count the probability of successfully identifying strong, WAI, WBC predictors using & the type-1 error using forward selection:

```{r}
set.seed(123)

n_sim <- 100
corr_vec <- seq(0.1, 0.9, 0.2)
varying_corr_out_forward <- data.frame(
  cor_val = rep(corr_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(corr_vec) * n_sim),
  pct_wai_vec = rep(NA, length(corr_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(corr_vec) * n_sim),
  type1_error_vec = rep(NA, length(corr_vec) * n_sim),
  power_vec = rep(NA, length(corr_vec) * n_sim)
)

for (j in 1:length(corr_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run forward selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run forward selection
    data_obj <- data_gen1(cor = corr_vec[j])
    forward_out <- fit_forward(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, forward_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_corr_out_forward[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_corr_out_forward[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_corr_out_forward[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_corr_out_forward[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_corr_out_forward[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

write.csv(varying_corr_out_forward, "./simulation_outputs/varying_corr_out_forward.csv")

```

### LASSO selection
Count the probability of successfully identifying strong, WAI, WBC predictors using & the type-1 error using LASSO:
```{r}
set.seed(123)

n_sim <- 100
corr_vec <- seq(0.1, 0.9, 0.2)

varying_corr_out_lasso <- data.frame(
  cor_val = rep(corr_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(corr_vec) * n_sim),
  pct_wai_vec = rep(NA, length(corr_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(corr_vec) * n_sim),
  type1_error_vec = rep(NA, length(corr_vec) * n_sim),
  power_vec = rep(NA, length(corr_vec) * n_sim)
)

for (j in 1:length(corr_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run LASSO selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run forward selection
    data_obj <- data_gen1(cor = corr_vec[j])
    lasso_out <- fit_lasso(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, lasso_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_corr_out_lasso[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_corr_out_lasso[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_corr_out_lasso[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_corr_out_lasso[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_corr_out_lasso[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

write.csv(varying_corr_out_lasso, "./simulation_outputs/varying_corr_out_lasso_223_1.csv")
```

## Varying the number of WBC predictors
### Forward 
```{r}
set.seed(123)

n_sim <- 100
wbc_vec <- c(2, 6, 10, 14)

varying_wbc_out_forward <- data.frame(
  wbc_num = rep(wbc_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(wbc_vec) * n_sim),
  pct_wai_vec = rep(NA, length(wbc_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(wbc_vec) * n_sim),
  type1_error_vec = rep(NA, length(wbc_vec) * n_sim),
  power_vec = rep(NA, length(wbc_vec) * n_sim)
)

for (j in 1:length(wbc_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run forward selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run LASSO
    data_obj <- data_gen1(n_wbc = wbc_vec[j])
    forward_out <- fit_forward(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, forward_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_wbc_out_forward[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_wbc_out_forward[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_wbc_out_forward[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_wbc_out_forward[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_wbc_out_forward[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

varying_wbc_out_forward
write.csv(varying_wbc_out_forward, "./simulation_outputs/varying_wbc_out_forward.csv")
```


### LASSO
```{r}
set.seed(123)

n_sim <- 100
wbc_vec <- c(2, 6, 10, 14)

varying_wbc_out_lasso <- data.frame(
  wbc_num = rep(wbc_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(wbc_vec) * n_sim),
  pct_wai_vec = rep(NA, length(wbc_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(wbc_vec) * n_sim),
  type1_error_vec = rep(NA, length(wbc_vec) * n_sim),
  power_vec = rep(NA, length(wbc_vec) * n_sim)
)

for (j in 1:length(wbc_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run LASSO selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run LASSO
    data_obj <- data_gen1(n_wbc = wbc_vec[j])
    lasso_out <- fit_lasso(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, lasso_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_wbc_out_lasso[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_wbc_out_lasso[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_wbc_out_lasso[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_wbc_out_lasso[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_wbc_out_lasso[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

write.csv(varying_wbc_out_lasso, "./simulation_outputs/varying_wbc_out_lasso.csv")
```

## Varying the number of WAI predictors
### Forward
```{r}
set.seed(123)

n_sim <- 100
wai_vec <- c(2, 6, 10, 14)

varying_wai_out_forward <- data.frame(
  wai_vec = rep(wai_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(wai_vec) * n_sim),
  pct_wai_vec = rep(NA, length(wai_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(wai_vec) * n_sim),
  type1_error_vec = rep(NA, length(wai_vec) * n_sim),
  power_vec = rep(NA, length(wai_vec) * n_sim)
)

for (j in 1:length(wai_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run forward selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run LASSO
    data_obj <- data_gen1(n_wai = wai_vec[j])
    forward_out <- fit_forward(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, forward_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_wai_out_forward[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_wai_out_forward[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_wai_out_forward[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_wai_out_forward[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_wai_out_forward[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

write.csv(varying_wai_out_forward, "./simulation_outputs/varying_wai_out_forward.csv")
```



### LASSO
```{r}
set.seed(123)

n_sim <- 100
wai_vec <- c(2, 6, 10, 14)

varying_wai_out_lasso <- data.frame(
  wai_vec = rep(wai_vec, each = n_sim),
  pct_strong_vec = rep(NA, length(wai_vec) * n_sim),
  pct_wai_vec = rep(NA, length(wai_vec) * n_sim),
  pct_wbc_vec = rep(NA, length(wai_vec) * n_sim),
  type1_error_vec = rep(NA, length(wai_vec) * n_sim),
  power_vec = rep(NA, length(wai_vec) * n_sim)
)

for (j in 1:length(wai_vec)){
  
  pct_strong_vec <- rep(NA, n_sim)
  pct_wai_vec <- rep(NA, n_sim)
  pct_wbc_vec <- rep(NA, n_sim)
  type1_error_vec <- rep(NA, n_sim)
  power_vec <- rep(NA, n_sim)

  # run LASSO selection simulation in default data_gen1 setting for 100 times
  for (i in 1:n_sim){
    #print(paste("current sim: ", i))
    # generate data and run LASSO
    data_obj <- data_gen1(n_wai = wai_vec[j])
    lasso_out <- fit_lasso(data_obj)
  
    # get percentages of strong/wai/wbc indicators correctly identified
    pct_out <- check_pct_identified(data_obj, lasso_out)
  
    # assign above results to a vector
    pct_strong_vec[i] <- pct_out$pct_strong_identified
    pct_wai_vec[i] <- pct_out$pct_wai_identified
    pct_wbc_vec[i] <- pct_out$pct_wbc_identified
    type1_error_vec[i] <- pct_out$type1_error
    power_vec[i] <- pct_out$power
  
  } 
  
  # combine to data frame
  varying_wai_out_lasso[((j-1)*n_sim+1):(j*n_sim), 2] <- pct_strong_vec
  varying_wai_out_lasso[((j-1)*n_sim+1):(j*n_sim), 3] <- pct_wai_vec
  varying_wai_out_lasso[((j-1)*n_sim+1):(j*n_sim), 4] <- pct_wbc_vec
  varying_wai_out_lasso[((j-1)*n_sim+1):(j*n_sim), 5] <- type1_error_vec
  varying_wai_out_lasso[((j-1)*n_sim+1):(j*n_sim), 6] <- power_vec
  
  print(paste("current j: ", j))
    
}

write.csv(varying_wai_out_lasso, "./simulation_outputs/varying_wai_out_lasso.csv")
```




## simulation of beta_sse
### Varying correlation (0.1, 0.3, 0.5, 0.7, 0.9)
```{r}
## forward AIC
simu_forward_betasse<-function(n_sim,c){
  beta_sse=rep(NA,n_sim)
  for (i in 1:n_sim) {
    print(i)
    data_obj<-data_gen1(cor = c)
    forward_out<-fit_forward(data_obj)
  
    sse_out<-check_beta_se(data_obj,forward_out)
    beta_sse[i]=sse_out$sse
  }
  ##bind results together
  sim_sse<-data.frame(cor=rep(c,n_sim),beta_sse)
  return(sim_sse)
}

## simu for 100 times forward
sse_0.1<-simu_forward_betasse(100,0.1)
sse_0.3<-simu_forward_betasse(100,0.3)
sse_0.5<-simu_forward_betasse(100,0.5)
sse_0.7<-simu_forward_betasse(100,0.7)
sse_0.9<-simu_forward_betasse(100,0.9)

sse_forward<-rbind(sse_0.1,sse_0.3,sse_0.5,sse_0.7,sse_0.9)
write.csv(sse_forward,"output_analysis_files/sse_forward.csv")

## lasso
simu_lasso_betasse<-function(n_sim,c){
  beta_sse_lasso=rep(NA,n_sim)
  for (i in 1:n_sim) {
    print(i)
    data_obj<-data_gen1(cor = c)
    lasso_out<-fit_lasso(data_obj)
  
    sse_out<-check_beta_se(data_obj,lasso_out)
    beta_sse_lasso[i]=sse_out$sse
  }
  ##bind results together
  sim_sse<-data.frame(cor=rep(c,n_sim),beta_sse_lasso)
  return(sim_sse)
}

#### simu for 100 times lasso
sse_0.1_lasso<-simu_forward_betasse(100,0.1)
sse_0.3_lasso<-simu_forward_betasse(100,0.3)
sse_0.5_lasso<-simu_forward_betasse(100,0.5)
sse_0.7_lasso<-simu_forward_betasse(100,0.7)
sse_0.9_lasso<-simu_forward_betasse(100,0.9)

sse_lasso<-rbind(sse_0.1_lasso,sse_0.3_lasso,sse_0.5_lasso,sse_0.7_lasso,sse_0.9_lasso)
write.csv(sse_lasso,"output_analysis_files/sse_lasso.csv")
```


## 
```{r}

```

