---
title: "output_analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(readr)
```

## Varying correlation between strong and WBC indicators

Load simulation output:
```{r}
forward_cor_simdata <- read_csv("./simulation_outputs/varying_corr_out_forward.csv")
lasso_cor_simdata <- read_csv("./simulation_outputs/varying_corr_out_lasso.csv")
```

Change the above data to long format for easier visualization
```{r}
forward_cor_simdata_cleaned <- forward_cor_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(cor_val = factor(cor_val)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  mutate(method = "forward")

forward_cor_simdata_cleaned

lasso_cor_simdata_cleaned <- lasso_cor_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(cor_val = factor(cor_val)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  #group_by(cor_val, metric) %>% 
  #summarize(mean_pct = mean(value))
  mutate(method = "lasso")

combined_cor_simdata <- rbind(forward_cor_simdata_cleaned, lasso_cor_simdata_cleaned)
combined_cor_simdata
```

Compare the probability of correctly selecting strong, wai, wbc indicators for forward selection and LASSO, respectively:
```{r}
# WBC
combined_cor_simdata %>% 
  filter(metric == "pct_wbc_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~cor_val, nrow = 1) +
  labs(title = "% WBC predictors selected for different correlations \n between strong and WBC predictors",
       y = "%")

# WAI
combined_cor_simdata %>% 
  filter(metric == "pct_wai_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~cor_val, nrow = 1) +
  labs(title = "% WAI predictors selected for different correlations \n between strong and WBC predictors",
       y = "%")

```

Compare the power and type-1 errors for forward selection and LASSO, respectively:
```{r}
# power
combined_cor_simdata %>% 
  filter(metric == "power_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~cor_val, nrow = 1) +
  labs(title = "Power (%) for different correlations between strong and WBC predictors",
       y = "power (%)")

# Type-1 error
combined_cor_simdata %>% 
  filter(metric == "type1_error_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~cor_val, nrow = 1) +
  labs(title = "Type-1 error rate (%) for different correlations \n between strong and WBC predictors",
       y = "Type-1 error rate (%)")

```

## Varying the number of WBC predictors
```{r}
lasso_wbc_simdata <- read_csv("./simulation_outputs/varying_wbc_out_lasso.csv")
forward_wbc_simdata <- read_csv("./simulation_outputs/varying_wbc_out_forward.csv")
```


```{r}
forward_wbc_simdata_cleaned <- forward_wbc_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(wbc_num = factor(wbc_num)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  mutate(method = "forward")

forward_wbc_simdata_cleaned

lasso_wbc_simdata_cleaned <- lasso_wbc_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(wbc_num = factor(wbc_num)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  #group_by(cor_val, metric) %>% 
  #summarize(mean_pct = mean(value))
  mutate(method = "lasso")

combined_wbc_simdata <- rbind(forward_wbc_simdata_cleaned, lasso_wbc_simdata_cleaned)
combined_wbc_simdata
```

```{r}
# power
combined_wbc_simdata %>% 
  filter(metric == "power_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~wbc_num, nrow = 1) +
  labs(title = "Power (%) for different number of WBC predictors",
       y = "power %")

# Type-1 error
combined_wbc_simdata %>% 
  filter(metric == "type1_error_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~wbc_num, nrow = 1) +
  labs(title = "Type-1 error rate (%) for different number of WBC predictors",
       y = "type-1 error %")
```

## Varying the number of WAI predictors
```{r}
lasso_wai_simdata <- read_csv("./simulation_outputs/varying_wai_out_lasso.csv")
forward_wai_simdata <- read_csv("./simulation_outputs/varying_wai_out_forward.csv")

forward_wai_simdata_cleaned <- forward_wai_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(wai_vec = factor(wai_vec)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  mutate(method = "forward")

#forward_wai_simdata_cleaned

lasso_wai_simdata_cleaned <- lasso_wai_simdata %>% 
  dplyr::select(-1) %>% 
  mutate(wai_vec = factor(wai_vec)) %>% 
  pivot_longer(cols = c(pct_strong_vec, pct_wai_vec, pct_wbc_vec, type1_error_vec, power_vec),
                           names_to = "metric") %>% 
  #group_by(cor_val, metric) %>% 
  #summarize(mean_pct = mean(value))
  mutate(method = "lasso")

combined_wai_simdata <- rbind(forward_wai_simdata_cleaned, lasso_wai_simdata_cleaned)
#combined_wbc_simdata

# power
combined_wai_simdata %>% 
  filter(metric == "power_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~wai_vec, nrow = 1) +
  labs(title = "Power (%) for different number of WAI predictors",
       y = "%")

# Type-1 error
combined_wai_simdata %>% 
  filter(metric == "type1_error_vec") %>% 
  ggplot(aes(x = method, y = value)) +
  geom_violin(alpha = 0.5, width = 1, aes(fill = method)) + 
  geom_boxplot(alpha = 0.5, width = 0.2) +
  #geom_jitter(alpha = 0.8, width = 0.2, aes(color = method)) +
  facet_wrap(~wai_vec, nrow = 1) +
  labs(title = "Type-1 error rate (%) for different number of WAI predictors",
       y = "%")

```







